---
title: "progress-report-kcjacobs13"
author: "Keenan Jacobs"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1: Scraping Hall of Fame Data for 2024

First, scrape the Hall of Fame voting data for 2024 from Baseball Reference.

```{r}
library(rvest)
library(dplyr)

# Define the URL
url <- "https://www.baseball-reference.com/awards/hof_2024.shtml"

# Read the HTML and extract the tables
html <- read_html(url)
tables <- html_table(html)

# Select the relevant table (assumes the first table is the correct one)
raw_data <- tables[[1]]

# Display the first few rows of the raw data
head(raw_data)

```

# Step 2: Clean the Data

The raw data does not have proper column names. Use the first row as the header and clean the dataset.

```{r}
# Extract the column names from the first row and clean the dataset
actual_col_names <- raw_data[1, ]
colnames(raw_data) <- actual_col_names
cleaned_data <- raw_data[-1, ]

# Display the first few rows of the cleaned data
head(cleaned_data)

```

Now we adjuct the variable types.

```{r}
library(readr)
# Make all column names unique
colnames(cleaned_data) <- make.unique(colnames(cleaned_data))

# Parse numeric columns, including Votes and %vote
cleaned_data <- cleaned_data %>%
  mutate(
    Votes = readr::parse_number(Votes),
    PercentVote = readr::parse_number(`%vote`)
  )

# Display the structure of the cleaned data
str(cleaned_data)



```

# Step 3: Merge with existing Data

Load the existing Hall of Fame dataset from the Lahman package and append the 2024 data.

```{r}
library(Lahman)

# Load the existing Hall of Fame data
hof_data <- HallOfFame

# Select and rename columns to match the existing Hall of Fame table
new_data <- cleaned_data %>%
  select(
    playerID = Name,
    yearID = `YoB`,
    votes = Votes,
    needed = `PercentVote`,
    inducted = `HOFm`
  ) %>%
  mutate(
    yearID = 2024,
    inducted = ifelse(inducted >= 75, "Y", "N"),
    category = "Player"
  )

# Combine datasets
extended_hof <- bind_rows(hof_data, new_data)

# Save the combined dataset to a CSV file
write.csv(extended_hof, file = "HallOfFame.csv", row.names = FALSE)

```

# Step 4: Data Visualization

For fun, create a plot showing the percentage of votes received by players over time.

```{r}
library(ggplot2)

# Plot the voting trends
ggplot(data = extended_hof, aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2024)) +
  labs(
    title = "Hall of Fame Voting Trends",
    x = "Year",
    y = "Number of Players",
    fill = "Inducted"
  )

```

The 2024 Hall of Fame voting data has been successfully scraped, cleaned, and integrated with the Lahman dataset. The expanded dataset has been saved as 'HallOfFame.csv'

